\documentclass[12pt,draft,letter]{article}
\usepackage{url}
\usepackage{natbib}
\usepackage{endnotes}
\title{Interactional Theories of Cities}
\author{Forest Gregg}

\let\footnote=\endnote
\let\cite=\citep


\usepackage{todonotes}

\begin{document}
\maketitle
\noindent This document has \makeatletter\@@input|"echo `texcount -1 -inc ecological_approaches.tex`| cut -c1-4"\makeatother words
\input{intro.tex}
\input{competition.tex}
\input{antagonism.tex}
\input{mutualism.tex}
\section*{Future Directions}
If mutualism is the prime driver of American urbanism, we have only
really begun to understand it's dynamics. Here are a number of
critical questions of whether and how mutualism will affect American
cities: how does mutualistic coherence affect development, where do
aesthetics come from, and what happens to people outside of the
mutualistic circle.

\section*{Detecting Coherence}
In the mutualistic theories of firms or people, there is an implicit
concept of a critical, coherent population. Above some threshold
number of complementary people or organizations, people see an area as
having a scene and can coordinate on that shared perception. Above
some threshold of complementary firms, there is industry cluster and
firms and can benefit from co-location. Below some threshold, there is
no there.

When a critical threshold is passed, the mutualistic feedback loop
should cause the proliferation of complementary populations. The death
rate should fall or the birth rate should increase or both. After
crossing a threshold, the concentration of complementary populations
should increase in proportion to the current population,
i.e. exponentially. Eventually, the forces of competition will limit
that growth either through price of land or the limits of client
demand.

Even if a scene or cluster is no longer growing, we should still
expect that the cluster will have different birth or survival rates
for complementary and noncomplementary populations compared to a
place with a different scene or cluster. Birth and survival also
include in and out migration from the place.

Critical coherence is both a interesting concept in it's own right and
a critical test for mutualistic theories. If we cannot detect expected
patterns of take-off after a certain threshold is reached, that will
be a serious blow to the ideas of industry clusters or scenes. If the
pattern does exist, then coherence thresholds will be a useful tool
for understanding the trajectory of cities and useful diagnostic for
local economic development.


\section*{Elective Affinities Between Post-Industrial Work and
  Aesthetics}

If people are attracted to places based upon aesthetic, lifestyle
considerations, then the origin of their aesthetics becomes a
critical question for urban sociology. In his discussion of how one's
environment acts on a person to form our aesthetic judgement, Bourdieu
is likely very helpful here. However, in his wider claims he argues
that at any particular time the space of taste should largely be
organized along two axes. This does not seem enough to describe the
many scenes that exist within and between cities. If Clark and
Silver's thesis is correct there are many more than two dimensions
along which the aesthetics of a place can vary.\footnote{Bourdieu is
  careful to not specify the aesthetic dimensions that taste vary
  upon, just that at any time the space of taste will be largely
  organized along two axes.\cite{bourdieu_distinction:_1986}}

Looking more closely at the connection between work and aesthetics
will likely be very useful for urban sociology. If there is an
elective affinity between different types of work and aesthetics, then
places with concentrations of particular types of work, industry
clusters, should express particular aesthetics, and vice versa. As
mutualistic feedback leads cities to specialize in different types of
industries, they should, as part of the same process, differentiate in
their aesthetics.

\subsection*{The Rest of Us}
In both the theories of industry clusters and scenes, the key actor
are very high skill workers who's labor is critical to post-industrial
production. These types of workers are minority of Americans and a
minority of city dwellers. How do the rest of us fair in a mutualistic
city, if we are not part of the virtuous cycle.

Saskia Sassen provides a partial answer. These highly paid workers
create substantial demand for low-skill service jobs. However, there
is a great more work to be done on how the political economy of places
where a coherent set of firms and workers generate most of the wealth,
but are not necessarily concentrated into large organizations. Should
we expect to see guild like behavior?

At the neighborhood level, many places don't have much of a
scene. Does living in a place without a distinct aesthetic coherence
have any predictable outcomes for the ability of a community to solve
problems or the welfare of its residents?



\newpage
\begingroup \parindent 0pt
\parskip 2ex
\def\enotesize{\normalsize}
\theendnotes
\endgroup

\newpage
\bibliographystyle{plainnat}
\bibliography{Ecology}


\end{document}
